---
title: "Exploratory Data Analysis: Fire Incident Dataset"
output: html_notebook
---


```{r, include=FALSE}

library(data.table)
library(tidyverse)
library(lubridate)
```



```{r, include = FALSE}
### Load Data

taxmap.data <- fread('data/tax_map.csv')                          ## From raw_with_match folder
taxrolls.data <- fread('data/Historic_Secured_Property_Tax_Rolls.csv') 

fire.incidents.data <- fread('data/matched_Fire_Incidents.csv')   ## From raw_with_matched folder

```

This is a test to see if there is a correlation between age of building and fire incidents in San Francisco. The analysis is based upon data tax data and fire incident data from the City of San Francisco. The fire data spans years 2011 through 2015. 

_Note 8-2-2017: The code below was run upon data that contained data entry errors. Consider re-running with cleaned-up data if available._ 

```{r}
### Data Processing

# Join tax data with fire incident data
# Use tax data from FY 2015
# Use fire incident data from 2011 through 2015

temp_tax <- taxrolls.data %>%
  inner_join(taxmap.data, by = c('Block and Lot Number' = 'Tax_ID'))

temp_tax_2015 <- filter(temp_tax, `Closed Roll Fiscal Year` == 2015)

temp_incident <- fire.incidents.data %>%
  mutate(`Incident Date` = ymd(`Incident Date`)) %>%
  filter(`Incident Date` >= "2011-01-01" & `Incident Date` <= "2015-12-31")

joined_incident <- temp_tax_2015 %>%
  left_join(temp_incident, by = c('EAS BaseID Matched' = 'EAS'))
```

```{r}
# Calculate count of incidents, group by Block and Lot Number and Year Built
# Assume age > 0 or age < 200 are outliers due to collection error

collapsed_incident <- joined_incident %>%
  group_by(`Block and Lot Number`, `Year Property Built`) %>%
  summarize(count_incident = sum(!is.na(`Incident Number`))) %>%
  mutate(age = 2017 - `Year Property Built`) %>%
  filter(age > 0 & age < 200)

# Scatterplot: x = age, y = count of incidents

ggplot(data = collapsed_incident, mapping = aes(x = age, y = count_incident)) + 
  geom_point(alpha = 1/2) + labs(x = 'Building Age (Age)', y = 'Count of fire incidents') + theme(plot.title = element_text(hjust = 0.5)) + ggtitle("Fire Incidents by Year Built")
```

```{r, include = FALSE}
ggplot(data = collapsed_incident, mapping = aes(x = age)) + geom_histogram(binwidth = 10) + ggtitle('Building Age Count') + theme(plot.title = element_text(hjust = 0.5))
```

Bin the ages of buildings, and then plot buildings that had one or more fire incidents.

```{r}

##  Unique EAS codes that have had >= 1 fire incidents in our time window 

EAS_buildings <- unique(temp_incident$EAS)

## Buildings with >= 1 fire incidents
buildings_w_fire <- temp_tax_2015 %>%
  mutate(fire = `EAS BaseID Matched` %in% EAS_buildings) %>%
  mutate(age = 2017 - `Year Property Built`) %>%
  mutate(decade = plyr::round_any(`Year Property Built`, 10, floor)) %>%
  group_by(decade) %>%
  summarize(proportion_fire = sum(fire)/ n(), count = n()) %>%
  filter(decade >= 1870 & decade <= 2010)

print(buildings_w_fire)

```
